Interisland movements of woodland caribou in Newfoundland

### 2020-02-19 paper/manuscript.Rmd
P3

Caribou were introduced to Fogo Island (n = 26 animals introduced between
1964–67) and Change Island (n = 5 animals introduced in 1964) from the main
Island of Newfoundland as part of a series of translocations and introductions
throughout the province (Bergerud & Mercer 1989). While caribou populations are
in decline throughout Canada (Festa-Bianchet et al. 2011), as well as in
Newfoundland (Bastille-Rousseau et al. 2013), the Fogo Island herd has not
declined to the same extent and currently consists of approximately 300 animals
(Newfoundland and Labrador Wildlife Division, unpublished data). Since 2016 we
have been studying the Fogo Island caribou herd (Bonar et al. 2017; Bonar et al.
2018; Peignier et al. 2019). To date, we have fitted XX adult female caribou
with GPS collars (1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario,
Canada), which were programmed to collect relocation fixes every 2 hours (see
Bonar et al. 2018 for details).

We posit two potential mechanisms explaining the
trade-offs associated with swimming in the North Atlantic Ocean: forage
limitation and the associated density-dependent habitat selection and predator
avoidance. For caribou living on small islands in the ocean, the same notions
for why caribou swim in freshwater may also explain why caribou swim in the
ocean (Leblond et al. 2016). Specifically, forage limitation (Miller 2002) and
predator avoidance (Jeffery et al. 2007) have been proposed as potential
explanations.

While caribou often swim across streams, rivers, and lakes (Leblond et al.
2016), few observations exist of caribou swimming in the ocean (see Table 1). Of
the existing or inferred observations of caribou swimming in the ocean, none
appear to be as prolific as the number and frequency of swimming events we
observed. We surmise that for caribou living on small islands, the primary
drivers of swimming are likely related to forage limitation and predator
avoidance.

>>>>

<!--DO WE NEED THIS SECTION?--> 
Caribou were introduced to Fogo Island (n = 26 animals introduced between
1964–67) and Change Island (n = 5 animals introduced in 1964) from the main
Island of Newfoundland as part of a series of translocations and introductions
throughout the province (Bergerud & Mercer 1989). While caribou populations are
in decline throughout Canada (Festa-Bianchet et al. 2011), as well as in
Newfoundland (Bastille-Rousseau et al. 2013), the Fogo Island herd has not
declined to the same extent and currently consists of approximately 300 animals
(Newfoundland and Labrador Wildlife Division, unpublished data). Since 2016 we
have been studying the Fogo Island caribou herd (Bonar et al. 2017; Bonar et al.
2018; Peignier et al. 2019). To date, we have fitted XX adult female caribou
with GPS collars (1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario,
Canada), which were programmed to collect relocation fixes every 2 hours (see
Bonar et al. 2018 for details).

We posit two potential mechanisms explaining the trade-offs associated with
swimming in the North Atlantic Ocean: forage limitation and predator avoidance.
For caribou living on small islands in the ocean, the same notions for why
caribou swim in freshwater may also explain why caribou swim in the ocean
(Leblond et al. 2016). Specifically, forage limitation (Miller 2002) and
predator avoidance (Jeffery et al. 2007) have been proposed as potential
explanations.
